Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2012-12-25 spie/_region_.tex
text doneish

\message{ !name(paper.tex) !offset(-42) }
\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}

\author{Peter Cheng, Michael Anderson, Stewart He, Avideh Zakhor
\skiplinehalf
University of California, Berkeley ; \\
}

%>>>> Further information about the authors, other than their 
%  institution and addresses, should be included as a footnote, 
%  which is facilitated by the \authorinfo{} command.

\authorinfo{Further author information: (Send correspondence to A.A.A.)\\A.A.A.: E-mail: aaa@tbk2.edu, Telephone: 1 505 123 1234\\  B.B.A.: E-mail: bba@cmp.com, Telephone: +33 (0)1 98 76 54 32}
%%>>>> when using amstex, you need to use @@ instead of @

\message{ !name(paper.tex) !offset(197) }

>>>>

\message{ !name(paper.tex) !offset(594) }
\subsection{Blending}
\label{sec:blending}
We now apply the same blending process to the two texturing methods:
tile caching or seam minimization.

Although the image alignment steps and image selection in both methods
attempt to minimize all mismatches between images, there are
occasional unavoidable discontinuities in the final texture due to
different lighting conditions or inaccuracies in planar geometry or
projection. These can however be treated and smoothed over by applying
alpha blending over image seams.  Whether the units we are blending
are rectangularly-cropped images or rectangular tiles, we can apply
the same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size. We have found $l_2 = \frac{l_1}{2}$ to provide a
good amount of blending without blurring features. For the seam
minimization method, we have already ensured overlap between
images. To enforce consistent blending however, we add a minimum
required overlap distance while solving the shortest path problem in
Section \ref{sec:seamMinimization}. Additionally, if images overlap in
a region greater than the overlap distance, we only apply blending
over an area equal to the overlap distance.

After blending pixels linearly across overlapping regions using alpha
blending, texture mapping is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the seam minimization approach still exhibits better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) has the best visual quality among the textures
in Figure \ref{fig:compareAll}.

\message{ !name(paper.tex) !offset(604) }

### 2012-12-25 spie/paper.tex
text doneish

The remainder of the paper is organized as follows. Section X explains
how input images are projected onto our geometry and simple approaches
towards texturing. Section X covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section X contains our approach towards image alignment,
followed by Section X, which describes compositing approaches. Section
X contains further results and conclusions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images. Given our high number of input images, one approach might be to more intelligently only select images that match up well, but the far more robust solution is to match up images directly such that we maximize our amount of useful images.

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger image is a
commonly performed task, with many successful approaches over the past
few decades. Parts of images are first matched to eachother, through
direct pixel comparisons, or more commonly through feature detection
and matching. Images are then adjusted to maximize matches, either by
directly calculating homographies between pairs of images, or by
modifying camera poses from 1 to 6 degrees of freedom.

Feature detection and matching works best when multiple unique visual
references exist in the environment and are present within multiple
images. This allows for accurate matching when texturing detailed
objects at high resolutions CiTATION, or when creating outdoor
panoramas CITATION. In contrast, our indoor environments have a high
prevalence of bare surfaces, as well as repeating textures that cause
difficulty in distinguishing features. The lack of strong reference
points results in high uncertainty when matching images together.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homographies, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this surface,
the mosaicing produces errors that cause straight lines to appear as
distorted, despite the fact that it was generated after careful hand
tuning. Many areas with fewer visual features simply failed outright
using this approach.

@@

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom. CITATION. Applying the approach in the LIU2010INDOOR paper
however, results again in error accumulation and drift as shown in
Figure \ref{fig:mosaic3D}(a). Furthermore, such nonlinear optimization
approaches have an extremely high complexity and runtime, and the
entire solution must be recomputed given any changes in input data.

@@

In this section, we describe our method of efficient and robust image
alignment. Our approach consists of four parts. First, all images are
projected onto the surface and rotated such that detected lines within
them are parallel to lines comprising the surface's boundary and
intersection with other surfaces. Next, these projections are shifted
such that interior lines near surface intersections are matched
up. Following this, occlusion checks are performed across each image
to remove invalid parts of each image for the target surface. These
first three steps are all image-independent and can be considered
pre-processing steps and performed in parallel. Finally, we detect
SIFT feature matches between pairs of images and solve a linear least
squares problem in 2D to maximize matches.

\subsection{Geometry-based Alignment}
\label{sec:projectionAndRotation}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we can detect
lines in the projections using Hough transforms. Experience and
intuition show that walls in indoor environments contain linear
features that are either horizontal or vertical, often corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. Though not as relevant in our datasets, this can be extended
such that lines are oriented with a wall's boundaries, for example in
areas with a slanted roof.

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, it makes sense to rotate and align this boundary such that
the visual boundary between two surfaces matches the physical boundary
in our digital model.

@@

\subsection{Image Occlusion}
get from github i guess i dunno. or check avz email

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than intensity-based alignment due to the differences
in lighting as well as possible occlusion among our images, both of
which feature alignment is less sensitive to \cite{lai1999robust,
  lowe1999object, mikolajczyk2005performance, szeliski2006image}.
SIFTGPU MORE CITATION SIFT matches determine $d^x$ and $d^y$ distances
between each pair of features for two images on the plane, though
these distances may not always be the same for different
features. Since indoor environments often contain repetitive features
such as floor tiles or doors, we need to ensure that SIFT-based
distances are reliable. In order to mitigate the effect of incorrect
matches and outliers, the RANSAC framework \cite{fischler1981random}
is used for a robust estimate of the optimal $d^x_{i,j}$ and
$d^y_{i,j}$ distances between two images $i$ and $j$. The RANSAC
framework handles the consensus-building machinery, and requires a
fitting function and a distance function. For this application, the
fitting function simply finds the average distance between matches in
a pair of images. The distance function for a pair of points is chosen
to be the difference between those points' SIFT match distance and the
average distance computed by the fitting function. We use a 10 pixel
outlier threshold, so that SIFT matches are labeled as outliers if
their horizontal or vertical distances are not within 10 pixels of the
average distance computed by the fitting function.

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]
If only relative distances between images are included then there is
no way to determine the absolute location of any of the images and the
matrix becomes rank deficient. To fix this we choose the first image
to serve as the anchor for the rest, meaning all the absolute
distances are based on its original location. This is done by adding a
row with a $1$ in the first column and the rest zeros.

Finally, the $N$-dimensional observation vector $\vec{\gamma}$ is
constructed using the SIFT-based distances generated earlier in the
RANSAC matching stage. The last element in the observation vector is
the location of the first image determined by its original noisy
localization, from \cite{chen2010indoor, liu2010indoor}. Thus
$\vec{\gamma}$ can be written as:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there is often a break in the chain of images,
meaning that no SIFT matches are found between one segment of the
plane and another. In this case we add rows to the $A$ matrix and
observations to the $\vec{\gamma}$ vector that contain the original
noisy $x$ and $y$ distance estimates from the original camera
poses. Another way to do this is to add rows for all neighboring pairs
of images and solve a weighted least squares problem where the SIFT
distances are given a higher weight i.e. 1, and the noisy distances
generated by the original camera poses \cite{chen2010indoor,
  liu2010indoor} are given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these rotated and shifted
images results in the significant improvements shown in Figure
\ref{fig:compareAll}(c), as compared to Figure
\ref{fig:compareAll}(b).

@@

@@

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted for wall and ceiling planes in Figures
\ref{fig:projectionAngles}(a) and \ref{fig:projectionAngles}(b)
respectively. These oblique camera angles for ceilings translate into
textures that span extremely large areas once projected, as shown in
Figure \ref{fig:projectionAngles}(b). Using the tile-based texture
mapping criteria from Figure \ref{fig:scoringFunction}, such
projections have highly varying scores depending on the location on
the plane, as shown in Figure \ref{fig:projectionAngles}(b). Thus, the
tiling approach in Section \ref{sec:simpleTextureMapping} is a
reasonable choice for texturing floors and ceilings, as it uses the
parts of image projections that are viable choices for their
respective plane locations, e.g. areas near point A in Figure
\ref{fig:projectionAngles}(b), but not near point B.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while texture mapping a given tile. By using the
same image across tile boundaries, we can eliminate a discontinuity
altogether. If this is not possible because a tile is not visible in
images used by neighboring tiles, using similar images across tile
boundaries also leads to less noticeable discontinuities.

Essentially a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a viable candidate,
before searching through the entire set of valid images. The first
subset of images is the images selected by adjacent tiles that have
already been textured. We must first check which of these images can
map to $t$, and then of those, we make a choice according to the
scoring function in Figure \ref{fig:scoringFunction}. Before reusing
this image, we ensure it meets the criteria $\alpha < 45^\circ$, in
order to be considered a viable image, with $\alpha$ as the camera
angle as shown in Figure \ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we check the
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar, allowing us to minimize any differences
in localization error as well as a parallax effects. Again, if no
viable image is found according to the same criteria, we search the
entire set of candidate images, selecting based on the same scoring
function from Figure \ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(b). As compared to Figure
\ref{fig:compareAll}(a), discontinuities are reduced overall, but the
amount of remaining seams suggests that despite our high volume of
images, image selection alone cannot produce seamless textures. Camera
matrices, or the image projections themselves, have to be adjusted in
order to reliably generate clean textures.

\subsection{Shortest}
For wall planes however, most images are taken from close distances
and more or less head-on angles, resulting in higher resolution
fronto-parallel projections. As a result, for each tile on a wall
plane, the scoring function of Figure \ref{fig:scoringFunction} is
relatively flat with respect to a large number of captured images, as
they are all more or less head on. Thus, the scoring function is less
significant for walls, and it is conceivable to use a different
texturing strategy for them so as to minimize visible seams in texture
mapped planes. This can be done by choosing the smallest possible set
of images that (a) covers the entire plane and (b) minimizes the
amount of visible seams between them. A straightforward cost function
that accomplishes the latter is the Sum of Squared Differences (SSD)
of pixels in overlapping regions between all pairs of images, after
they have been rotated as described in Section
\ref{sec:projectionAndRotation}. Minimizing this cost function
encourages image boundaries to occur either in featureless areas, such
as bare walls, or in areas where images match extremely well.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the plane, using other polygons of
arbitrary geometry i.e. image projections, with the added constraint
of minimizing the cost function between chosen images.  This is a
complex problem, though we can take a number of steps to simplify it.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. To address this, we solve for
a new shortest path after modifying the edges chosen by the original
shortest path solution such that they have higher cost than all other
edges in the DAG, and solve for a new shortest path. This ensures that
these edges are not chosen again unless they are the only available
option. Our second shortest path solution results in a new set of
chosen images, of which we only retain those that cover areas not
covered by the first set. This process can be repeated as many times
as needed, until holes are no longer present. This method is not as
optimal as a 2D-coverage solution would be, but it is a fast
approximation, and adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the
posters in the images, which have clear misalignment over seams in
Figure \ref{fig:compareAll}(c), but are much more aligned in Figure
\ref{fig:compareAll}(d). This is because the seam minimization
approach directly reduces the cost of each image boundary, while the
tile caching method uses a scoring function that only approximates
this effect. Furthermore, seam minimization guarantees the best
selection of images, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. The seam minimization approach
is also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images\footnote {For the seam minimization
  approach, occlusion checks are performed over the entirety of each
  image to determine available areas for texture mapping. Since indoor
  environments mostly consist of vertical or horizontal surfaces with
  high amounts of right angles, image projections remain largely
  rectangular after masking out occluded areas. Therefore, as a
  preprocessing step, we can recursively split each image into
  rectangular pieces and check for occlusions in a similar manner as
  in the tiling process.  }.

@@

\subsection{Blending}
\label{sec:blending}
We now apply the same blending process to the two texturing methods:
image alignment followed by either tile caching or seam minimization.

Although the preprocessing steps and image selection in both methods
attempt to minimize all mismatches between images, there are
occasional unavoidable discontinuities in the final texture due to
different lighting conditions or inaccuracies in planar geometry or
projection. These can however be treated and smoothed over by applying
alpha blending over image seams.  Whether the units we are blending
are rectangularly-cropped images or rectangular tiles, we can apply
the same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size. For the seam minimization method, we have
already ensured overlap between images. To enforce consistent blending
however, we add a minimum required overlap distance while solving the
shortest path problem in Section
\ref{sec:seamMinimization}. Additionally, if images overlap in a
region greater than the overlap distance, we only apply blending over
an area equal to the overlap distance.

After blending pixels linearly across overlapping regions using alpha
blending, texture mapping is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the seam minimization approach still exhibits better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) has the best visual quality among the textures
in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}

>>>>

The remainder of the paper is organized as follows. Section X explains
how input images are projected onto our geometry and simple approaches
towards texturing. Section X covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section X contains our approach towards image alignment,
followed by Section X, which describes compositing approaches. Section
X contains further results and conclusions, as well as areas for
future work.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images. Given our high number of input images, it
makese sense to intelligently select only the images that match up
well, as performed in Section X, but for our solution to be robust, we
must align images directly such that we maximize the accuracy of each image.

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger image is a
commonly performed task, with many successful approaches over the past
few decades. Parts of images are first matched to eachother, through
direct pixel comparisons, or more commonly through feature detection
and matching. Images are then adjusted to maximize matches, either by
calculating homographies between pairs of images, or by
modifying camera poses from 1 to 6 degrees of freedom.

Feature detection and matching works best when multiple unique visual
references exist in the environment and are present within multiple
images. In contrast, our indoor environments have a high
prevalence of bare surfaces, as well as repeating textures that cause
difficulty in disambiguating features. This lack of strong reference
points results in high uncertainty when matching images together.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homographies, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this surface,
the mosaicing produces errors that cause straight lines to appear distorted, despite the fact that it was generated after careful hand
tuning. Many areas with fewer visual features simply failed outright
using this approach.

@@

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom. CITATION. Applying the approach in the LIU2010INDOOR paper
however, results again in error accumulation and drift as shown in
Figure \ref{fig:mosaic3D}(a). Furthermore, such nonlinear optimization
approaches IS IT? BETTER CHECK have high complexity and runtime, and the
entire solution must be recomputed given any changes in input data.

@@

In this section, we describe our method of efficient and robust image
alignment. Our approach consists of four parts. First, all images are
projected onto the surface and rotated such that detected lines within
them are parallel to lines comprising the surface's boundary and
intersection with other surfaces. Next, these projections are shifted
such that interior lines near surface intersections are matched
up. Following this, occlusion checks are performed to remove invalid
parts of each image for the target surface. These first three steps
are all image-independent and are performed in parallel. Finally, we
detect SIFT feature matches between pairs of images and solve a linear
least squares problem in 2D to maximize matches.

EVERYTHING IS IN 2D BECAUSE PITCH BECOMES WHATEVO

\subsection{Geometry-based Alignment}
\label{sec:projectionAndRotation}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we can detect
lines in the projections using Hough transforms. Experience and
intuition show that walls in indoor environments contain linear
features that are either horizontal or vertical, often corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. Though not relevant in our datasets, this can be extrapolated
such that lines are oriented with a wall's boundaries, for example in
areas with a slanted roof.

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, the image is rotated and shifted in 2D such that
the visual boundary between two surfaces matches the physical boundary
in our digital model.

@@

\subsection{Image Occlusion}
get from github i guess i dunno. or check avz email. maybe put in appendix

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lai1999robust, lowe1999object, mikolajczyk2005performance,
  szeliski2006image}.  SiftGPU, by PERSON CITATION, is used both for
feature detection as well as pairwise matching. These matches
determine $d^x$ and $d^y$ distances between each pair of features for
two image projections, though these distances may not always be the
same for different features. Since indoor environments often contain
repetitive features such as floor tiles or doors, we need to ensure
that SIFT-based distances are reliable. In order to mitigate the
effect of incorrect matches and outliers, the RANSAC framework
\cite{fischler1981random} is used for a robust estimate of the optimal
$d^x_{i,j}$ and $d^y_{i,j}$ distances between two images $i$ and
$j$. The RANSAC framework handles the consensus-building machinery,
and requires a fitting function and a distance function. For this
application, the fitting function simply finds the average distance
between matches in a pair of images. The distance function for a pair
of points is chosen to be the difference between those points' SIFT
match distance and the average distance computed by the fitting
function. We use a 10 pixel outlier threshold, so that SIFT matches
are labeled as outliers if their horizontal or vertical distances are
not within 10 pixels of the average distance computed by the fitting
function.

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]
If only relative distances between images are included then there is
no way to determine the absolute location of any of the images and the
matrix becomes rank deficient. From Section X, a number of images were
anchored to geometry points, and thus their locations can be used to
anchor the rest in place. In case no anchor images exist, we simply
arbitrarily pick an image. In the above matrix, the first image is set
to be such an anchor.

Finally, the $N$-dimensional observation vector $\vec{\gamma}$ is
constructed using the SIFT-based distances generated in the RANSAC
matching stage. Elements in the observation vector corresponding to
anchor images are simply their locations as determined by the original
noisy localization. Thus $\vec{\gamma}$ can be written as:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there is often a break in the chain of images,
meaning that no SIFT matches are found between one segment of the
plane and another. To solve this, we add rows to the $A$ matrix and
observations to the $\vec{\gamma}$ vector corresponding to the
original noisy distances. We then solve a weighted least squares
problem where the SIFT distances and anchor values are given a high
weight e.g. 1, while the noisy distances are given a smaller weight
e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section X results in X, which is far more seamless, but a more
intelligent image selection scheme can still be used.

@@

In this section, we revisit the tile-based texturing approach from
Section X, with an added caching mechanism to reduce image
boundaries. This method works well given all manner of camera poses and
surfaces, but for cases where we have large sections of usable texture
from images, we propose an alternate method that further reduces image
boundaries.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section X, we saw that discontinuities occur where adjacent tiles
select different images. Though image alignment reduces or removes
many such discontinuities, there are still cases where seams are
visible, for example due to lighting differences or imprecise
matching. To reduce the cases where this happens, it makes sense to
take into account image selections made by neighboring tiles while
texture mapping a given tile. By using the same image across tile
boundaries, we can eliminate a discontinuity altogether. If this is
not possible because a tile is not visible in images used by
neighboring tiles, using similar images across tile boundaries also
leads to less noticeable discontinuities.

Essentially a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a viable candidate,
before searching through the entire set of valid images. The first
subset of images is the images selected by adjacent tiles that have
already been textured. We must first check which of these images can
map to $t$, and then of those, we make a choice according to the
scoring function in Figure \ref{fig:scoringFunction}. Before reusing
this image, we ensure it meets the criteria $\alpha < 45^\circ$, in
order to ensure a high resolution projection, with $\alpha$ as the
camera angle as shown in Figure \ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we check the
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar, and thus likely matched more
accurately. Again, if no viable image is found according to the same
criteria, we search the entire set of candidate images, selecting
based on the same scoring function from Figure
\ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure X. Seams are now greatly reduced, and LIFE IS HAPPY.

@@

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted for wall and ceiling planes in Figures
\ref{fig:projectionAngles}(a) and \ref{fig:projectionAngles}(b)
respectively. These oblique camera angles for ceilings translate into
textures that span extremely large areas once projected, as shown in
Figure \ref{fig:projectionAngles}(b). Using the tile-based texture
mapping criteria from Figure \ref{fig:scoringFunction}, such
projections have highly varying scores depending on the location of a tile on
the plane, as shown again in Figure \ref{fig:projectionAngles}(b). Thus, the
tiling approach in Section \ref{sec:simpleTextureMapping} is a
good choice for texturing floors and ceilings, as it uses the
parts of image projections that maximize resolution and accuracy for their
respective plane locations, e.g. areas near point A in Figure
\ref{fig:projectionAngles}(b), but not near point B.

\subsection{Shortest Path Texturing}

For wall planes however, most images are taken from close distances
and head-on angles, resulting in higher resolution fronto-parallel
projections. As a result, for each tile on a wall plane, the scoring
function of Figure \ref{fig:scoringFunction} is relatively flat with
respect to a large number of captured images, as they are all more or
less head on. Thus, the scoring function is less significant for
walls, and it is conceivable to use a different texturing strategy for
them so as to directly minimize visible seams when texturing
them. This can be done by choosing the smallest possible set of images
that (a) covers the entire plane and (b) minimizes the amount of
visible seams between them. A straightforward cost function that
accomplishes the latter is the Sum of Squared Differences (SSD) of
pixels in overlapping regions between all pairs of images. Minimizing
this cost function encourages image boundaries to occur either in
featureless areas, such as bare walls, or in areas where images match
extremely well.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the planar surface, using other
polygons of arbitrary geometry i.e. image projections, with the added
constraint of minimizing the cost function between chosen images.
This is a complex problem, though we can take a number of steps to
simplify it.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. Since these holes are rare,
and generally fairly small, we use a greedy approach, repeatedly
filling the hole with images that result in the lowest SSD costs in a
10 px border region around the hole. This method is not as optimal as
a 2D-coverage solution would be, but it is a fast approximation, and
adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the
posters in the images, which have clear misalignment over seams in
Figure \ref{fig:compareAll}(c), but are much more aligned in Figure
\ref{fig:compareAll}(d). This is because the seam minimization
approach directly reduces the cost of each image boundary, while the
tile caching method uses a scoring function that only approximates
this effect. Furthermore, seam minimization guarantees the best
selection of images, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. The seam minimization approach
is also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images.

@@

\subsection{Blending}
\label{sec:blending}
We now apply the same blending process to the two texturing methods:
tile caching or seam minimization.

Although the image alignment steps and image selection in both methods
attempt to minimize all mismatches between images, there are
occasional unavoidable discontinuities in the final texture due to
different lighting conditions or inaccuracies in planar geometry or
projection. These can however be treated and smoothed over by applying
alpha blending over image seams.  Whether the units we are blending
are rectangularly-cropped images or rectangular tiles, we can apply
the same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size.  We have found $l_2 = \frac{l_1}{2}$ to
provide enough blending without blurring features. For the seam
minimization method, we have already ensured overlap between
images. To enforce consistent blending however, we add a minimum
required overlap distance of 40 px while solving the shortest path
problem in Section \ref{sec:seamMinimization}. Additionally, if images
overlap in a region greater than the overlap distance, we only apply
blending over an area equal to the overlap distance.

After blending pixels linearly across overlapping regions using alpha
blending, our texture mapping process is complete. Figures
\ref{fig:compareAll}(e) and \ref{fig:compareAll}(f) show the blended
versions of Figures \ref{fig:compareAll}(c) and
\ref{fig:compareAll}(d) respectively. It is clear that the seam
minimization approach exhibits better alignment and fewer seams
than the tile-caching method, as Figure \ref{fig:compareAll}(f) has
the best visual quality among the textures in Figure
\ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the seam minimization approach, are
displayed in Figure \ref{fig:results}. High resolution texture
comparisons, as well as a walkthrough of a fully textured 3D model are
available in the accompanying video to this paper.

As mentioned earlier, our approach is fairly efficient, with the X by
X pixel image in figure X, spanning an X meter long wall and
consisting of X input images taking under a minute to generate on a
2.8Ghz dual-core laptop. While not fast enough for real-time
visualization, the procedure is quick enough for quick evaluation given
changes in various internal parameters or modifications to input data.

\subsection{Future stuff}
For the sake of simplicity, - cases where multiple whatever
are few, but for true generality, should go to 3d adjustments
